Mergeable Types

### 2017-02-27 paper/model.tex
fix wrapfig
minor edits: missing parens and spelling fix

\name requires a three-way \C{merge} function to merge concurrent
versions of a drawing canvas (see Fig.~\ref{fig:merge-canvas}) that
includes includes two concurrent versions (\C{v1} and \C{v2}),
and their least common ancestor (\C{lca}) - the version from which the
two concurrent versions evolved independently. The merge function can
\begin{wrapfigure}{l}{.4\textwidth}
\centering
\includegraphics[scale=0.43]{Figures/canvas-merging}
\caption{Merging concurrent versions (\C{v1} and \C{v2}) of a drawing
canvas. \C{lca} is their common ancestor.}
\label{fig:canvas-merging}
\end{wrapfigure}
make use of the pixel values of the common ancestor to merge the pixel
values on both the canvases. For instance, if the color of a pixel in
\C{v1} is white, and in \C{v2} it is green, and its color in \C{lca}
is white, then it means that only \C{v2} modified the color. Hence the
pixel is colored green in the merged canvas. On the other hand, if the
pixel is red in \C{v1}, then it means that both \C{v1} and \C{v2} have
modified the color. In such case, an appopriate color-mixing algorithm
can be used to determine the color of pixel.  For instance, the pixel
can be colored yellow - an additive combination of red and green. The
logic is illustrated in Fig.~\ref{fig:canvas-merging}.
\begin{figure}
\begin{center}
  \begin{ocaml}
let color_mix px1 px2 : pixel =
let f = Char.code in
let h x y = Char.chr @@ (x + y)/ 2 in
let (r1,g1,b1) = (f px1.r, f px1.g, f px1.b) in
let (r2,g2,b2) = (f px2.r, f px2.g, f px2.b) in
let (r,g,b) = (h r1 r2, h g1 g2, h b1 b2) in {r=r; g=g; b=b}

@@

The \name programming model lets programmers define and compose
concurrent computations around versioned data structures.
Fig.~\ref{fig:dali-monad} shows the signature of the \name module that
implements the programming model along the lines of the well-known
\C{State} monad~\cite{wadler-monad}. The monad encapsulates a
versioned functional state (\C{'a}) and the type \C{('a, 'b) t}
represents a monadic computation that returns a \C{'b} result.
Functions \C{return} and \C{bind} have their usual monadic
\begin{figure}
\begin{center}
  \begin{ocaml}
  module type DALI = sig
    type ('a, 'b) t
    val return : 'b -> ('a, 'b) t
    val bind : ('a, 'b) t -> ('b -> ('a, 'c) t) -> ('a, 'c) t
    val get_current_version: unit -> ('a, 'a) t
    val with_init_version_do: 'a -> ('a, 'b) t -> 'b
    val fork_version : ('a, 'b) t -> 'a unit t
    val sync_next_version: unit -> 'a -> ('a, 'a) t
  end
  \end{ocaml}
  \caption{Signature of the \name monad}
  \label{fig:dali-monad}
\end{center}
\end{figure}
interpretation. \C{get\_current\_version} is like the \C{State}
monad's \C{get}; it returns the versioned state encapsulated by the
monad. To initiate a computation, we use \C{with\_init\_version\_do}
which runs a monadic computation against a given initial version and
returns the result. The \C{fork\_version} operation returns a
computation that forks a new concurrent version from the current
version, and runs the given monadic computation asynchronously against
the forked version.  \C{sync\_next\_version} (simply called \C{sync})
accepts a \emph{proposal} for the next version of the state; this
proposal is the current local version of the state that reflects local
modifications not yet witnessed by any other concurrently executing
computation.  The operation returns (via a monad) the actual next
version, which becomes the current version for the rest of the
computation.  This version is created by merging the proposal with a
subset of concurrent versions that have become available since the
last merge or fork. Thus, \C{sync} effectively lets a computation
synchronize with a subset of concurrent computations to obtain their
latest updates.

>>>>

\name requires a three-way \C{merge} function to merge concurrent
versions of a drawing canvas (see Fig.~\ref{fig:merge-canvas}) that
includes includes two concurrent versions (\C{v1} and \C{v2}),
and their least common ancestor (\C{lca}) - the version from which the
two concurrent versions evolved independently. The merge function can
\begin{wrapfigure}{L}{.4\textwidth}
\centering
\includegraphics[scale=0.43]{Figures/canvas-merging}
\caption{Merging concurrent versions (\C{v1} and \C{v2}) of a drawing
canvas. \C{lca} is their common ancestor.}
\label{fig:canvas-merging}
\end{wrapfigure}
make use of the pixel values of the common ancestor to merge the pixel
values on both the canvases. For instance, if the color of a pixel in
\C{v1} is white, and in \C{v2} it is green, and its color in \C{lca}
is white, then it means that only \C{v2} modified the color. Hence the
pixel is colored green in the merged canvas. On the other hand, if the
pixel is red in \C{v1}, then it means that both \C{v1} and \C{v2} have
modified the color. In such case, an appopriate color-mixing algorithm
can be used to determine the color of pixel.  For instance, the pixel
can be colored yellow - an additive combination of red and green. The
logic is illustrated in Fig.~\ref{fig:canvas-merging}.
\begin{figure}
\begin{center}
  \begin{ocaml}
let color_mix px1 px2 : pixel =
let f = Char.code in
let h x y = Char.chr @@ (x + y)/ 2 in
let (r1,g1,b1) = (f px1.r, f px1.g, f px1.b) in
let (r2,g2,b2) = (f px2.r, f px2.g, f px2.b) in
let (r,g,b) = (h r1 r2, h g1 g2, h b1 b2) in {r=r; g=g; b=b}

@@

The \name programming model lets programmers define and compose
concurrent computations around versioned data structures.
Fig.~\ref{fig:dali-monad} shows the signature of the \name module that
implements the programming model along the lines of the well-known
\C{State} monad~\cite{wadler-monad}. The monad encapsulates a
versioned functional state (\C{'a}) and the type \C{('a, 'b) t}
represents a monadic computation that returns a \C{'b} result.
Functions \C{return} and \C{bind} have their usual monadic
\begin{figure}
\begin{center}
  \begin{ocaml}
  module type DALI = sig
    type ('a, 'b) t
    val return : 'b -> ('a, 'b) t
    val bind : ('a, 'b) t -> ('b -> ('a, 'c) t) -> ('a, 'c) t
    val get_current_version: unit -> ('a, 'a) t
    val with_init_version_do: 'a -> ('a, 'b) t -> 'b
    val fork_version : ('a, 'b) t -> ('a, unit) t
    val sync_next_version: unit -> 'a -> ('a, 'a) t
  end
  \end{ocaml}
  \caption{Signature of the \name monad}
  \label{fig:dali-monad}
\end{center}
\end{figure}
interpretation. \C{get\_current\_version} is like the \C{State}
monad's \C{get}; it returns the versioned state encapsulated by the
monad. To initiate a computation, we use \C{with\_init\_version\_do}
which runs a monadic computation against a given initial version and
returns the result. The \C{fork\_version} operation returns a
computation that forks a new concurrent version from the current
version, and runs the given monadic computation asynchronously against
the forked version.  \C{sync\_next\_version} (simply called \C{sync})
accepts a \emph{proposal} for the next version of the state; this
proposal is the current local version of the state that reflects local
modifications not yet witnessed by any other concurrently executing
computation.  The operation returns (via a monad) the actual next
version, which becomes the current version for the rest of the
computation.  This version is created by merging the proposal with a
subset of concurrent versions that have become available since the
last merge or fork. Thus, \C{sync} effectively lets a computation
synchronize with a subset of concurrent computations to obtain their
latest updates.
